Multi-Agent Pathfinding with Matching using Increasing Cost Tree Search

### 2021-06-27 docs/paper.tex
Removed 1 and 2 agents from three team plot, qualified statement about child pruning

	\section{Introduction}
	Finding a shortest path between two vertices on a graph is a fundamental problem that appears in many contexts like GPS navigation \cite{sturtevant2010} and network routing \cite{broch1998}. In some applications like autonomous aircraft towing \cite{morris2016} and train unit routing in shunting yards \cite{mulderij2020}, planning agents separately can result in conflicts between agents as they follow their paths. To prevent such conflicts, agents have to be planned together to find a combination of non-conflicting paths. The problem of finding such a combination is the Multi-Agent Pathfinding (MAPF) problem \cite{stern2019}. In finding optimal solutions, one of two objectives is typically optimized: the makespan which is the cost of the longest path, or the Sum of Individual path Costs (SIC). Algorithms for optimally solving MAPF include Increasing Cost Tree Search \cite{sharon2013}, Conflict-based search \cite{sharon2015} and Branch-and-Cut-and-Price \cite{lam2019}.

@@

	\section{Multi-agent Pathfinding with Matching} % Problem description
	\label{mapfm}
	\begin{wrapfigure}[19]{r}{.25\textwidth}
		\begin{minipage}{\linewidth}
			\centering\captionsetup[subfigure]{justification=centering}
			\includegraphics[width=\linewidth]{img/vertex-conflict}
			\subcaption{Vertex conflict}
			\label{fig:conflictsa}\par\vfill
			\includegraphics[width=0.75\linewidth]{img/edge-conflict}
			\subcaption{Edge conflict}
			\label{fig:conflictsb}
		\end{minipage}
		\vspace{-10pt}
		\caption{The two types of conflict (from Fig. 1 in \cite{stern2019})}\label{fig:conflicts}
	\end{wrapfigure}

@@

	\subsubsection{MDD generation for matching}
	To generate MDDs for MAPFM, each agent also has to consider not one goal but all matching goals. To do this, the MDD construction following the breadth-first search has to be changed: each goal becomes an endpoint from which paths to the start node are traced, resulting in an MDD with one start node and one or more terminal nodes.
	\begin{wrapfigure}{R}{.35\textwidth}
		\centering
		\includegraphics[width=\linewidth]{img/path}
		\caption{Pathological case for ICTS-m}
		\label{fig:path}
	\end{wrapfigure}
	\subsubsection{Pathological case}
	\label{pathology}
	As a consequence of the $\min$ operation in the root definition, a pathological case can be derived. Consider a team of 4 agents $a_1^1$ through $a_4^1$. Let them be arranged in a configuration as depicted in Figure \ref{fig:path}. All four agents match the color of the center goal and hence for each $i$, $c^*_i = 1$, so that the root is $(1,1,1,1)$ with SIC $4$. The optimal solution however is $(1,3,3,3)$ with SIC $10$ numbering the agents clockwise, meaning that $\Delta = 6$. It can be seen that instances of this form with arbitrarily large $\Delta$ can be constructed. As ICTS-m being a generalization of ICTS is $\mathcal{O}(k^\Delta)$, even just extending the 'limbs' of the instance to yield an ICT root of $(1,9,9,9)$ with $\Delta = 24$ will result in an unfathomably large number of nodes being evaluated assuming no pruning or other modifications to the base algorithm.
	%\subsubsection{Inverted MDD search}
	%The uniqueness of goals guarantees the existence of a goal node without vertex conflicts in searching MDDs in MAPF. In MAPFM, this is not the case and causes joint MDDs to be searched fully without finding a result. In addition, even if there exists a goal node, many of the paths represented by the joint MDD lead to conflicting maximal-depth nodes.

@@

	\restoregeometry
	\subsection{Experimental setup}
	Because MAPFM lies on the fringe of pathfinding research, no standard benchmark exists. For MAPF, the Moving AI benchmarks \cite{sturtevant2012} are often used and potentially a new MAPFM standard benchmark could be defined by generating suitable team-partitions of agents for these MAPF instances; however, it was noticed that optimally solving even small MAPFM instances with comparatively few agents and short distances to goals was already a challenge computationally. This observation informed the creation of a map generator by the author of \cite{jong2021} with various parameters.

@@

	\begin{enumerate}
		\item 25\% wall: a grid with 25\% of tiles being obstacles.
		\item 75\% wall: a grid with 75\% of tiles being obstacles. 
	\end{enumerate}
	For each of these types, maps were generated with between $1$ and $25$ agents either all in the same team or partitioned into three teams as evenly as possible. When solving the instances, for each combination of map-type and number of teams, the number of agents was increased until 0\% of instances could be solved within a timeout of 120s, at which point no more data was recorded for greater numbers of agents. This approach is similar to the method described in \cite{stern2019}.

@@

	\subsection{Experiment 1: ICTS-based MAPFM algorithms}
	To investigate the relative performance of various algorithms discussed in Section \ref{section:icts-matching}, a selection of both exhaustive and ICTS-m variants was made. Their performance in terms of fraction of problems solved within a timeout of 120s is shown in Figure \ref{fig:i-probs}. The meaning of the abbreviations used in the legend are as follows:
	\begin{itemize}
		\begin{minipage}{0.45\linewidth}
			\item S: Simple pruning
			\item E: Enhanced pruning
			\item C: Child pruning
		\end{minipage}
		\begin{minipage}{0.45\linewidth}
			\item ID: Independence detection
			\item B: Bounded search
			\item O: Approximate ordered enumeration
		\end{minipage}
	\end{itemize}
	Across scenarios, it is seen that typically exhaustive methods outperform ICTS-m algorithms (\ref{fig:i-probs}), with performance being more similar in the maps with 75\% wall. In particular on 75\% wall maps with three teams (\ref{fig:i-75-3-p}), the performance of ICTS-m+ID (and optional pruning) is seen to outperform bounded exhaustive search as $k$ increases, and some of these also manage to solve a small fraction of 75\% maps with one team that exhaustive methods cannot within the time-out (\ref{fig:i-75-1-p}).

	Between exhaustive variants, with exception of the ID variant, a type of hierarchy can be distinguished, best illustrated by \ref{fig:i-25-3-p}; almost without exception, Exh+S+C+B+O outperforms Exh+E+B+O which outperforms Exh+E+B which in turn outperforms Exh+E. Bounded search outperforms unbounded search as expected, given that the former does at most as much work as the latter and often less. Approximate ordering outperforming unordered enumeration is not quite as obvious from a theoretical standpoint, as discussed in \ref{ordered-enum}, and simply indicates that sorting works well in practice. Simple pruning with child pruning outperforming exhaustive pruning shows the potential of this pruning method. In particular, as $k$ grows larger and in scenarios with more obstacles (\ref{fig:i-75-1-p},\ref{fig:i-75-3-p}), it appears that having to search a smaller ICT is more important in terms of performance than being able to skip the $k$-agent low-level search more often. As discussed in Section \ref{future}, more work is needed to determine if such an improvement is to be expected more generally.

	The benefit of using ID together with ordered enumeration is less clear. Ordered enumeration with ID performs much better in \ref{fig:i-75-3-p}, a setting with more obstacles and more teams, but ordered enumeration without ID on the other hand is superior in an open setting with a single team (\ref{fig:i-25-1-p}). The overhead associated with the various planning and merging steps (see Algorithm \ref{id-algo}) could be the cause the slowdown observed once $M$ grows large (\ref{fig:i-25-1-p},\ref{fig:i-75-1-p}); once the upper bound is tightened, ICTs only have to be searched shallowly so there is no point in going through the steps of ID. This explains why no speedup is seen, as between ICTS-m and ICTS-m+ID. Another factor is that if all agents belong to the same team, it is unlikely that there will a significant benefit to using ID, as agents from the same team planned independently often route to the same goal.

@@

	CBM solving all problems with one team (Figure \ref{fig:r-25-1-p} and \ref{fig:r-75-1-p}), problems equivalent to anonymous MAPF instances, is at first surprising but follows directly from the definition of the CBM method. CBM solves single teams as anonymous MAPF instances in polynomial time, meaning that solving a single team reduces to a polynomial-time procedure, namely SSP. Following the results reported in \cite{ma2016}, it is expected that as the number of teams approaches $k$, CBM performance will deteriorate and thus become more similar to the results for exhaustive methods here reported.
	%	Still, some implementations might be more optimized than others so this might be a factor in performance characteristics and in particular in relative performance differences.
	\section{Conclusions and Future work}
	\label{conclusions}
	\label{future}
	The goal of this research was to seek methods for solving MAPFM based on ICTS. In this, the research has succeeded, but the scalability of the algorithms leaves something to be desired. Exhaustive methods, like those discussed in detail in Subsection \ref{exhaustive}, cannot be expected to scale well as teams grow larger due to their complexity, no matter what optimizations are put in place. Nonetheless, as was shown in Section \ref{experiments}, the more advanced exhaustive ICTS algorithms consistently outperform the ICTS-m based algorithms. Compared to the SIC-adaptation of the CBM method \cite{baauw2021}, it was seen that exhaustive methods generally perform relatively poorly if there are fewer teams and fewer obstacles, while being competitive once teams grow smaller and maps become more constrained. Delineating under which conditions exhaustive methods outperform CBM and vice versa is something for further research.

	%\section{Future work}
	A key idea for future research to look into is how to exploit the (inclusive) lower-bound on solution SIC that can be found using the Hungarian Algorithm \cite{kuhn55} or similar methods like the JVC algorithm \cite{jonker1987}. This lower-bound could be used in a branch-and-bound \cite{land1960} search of matchings. An algorithm for MAPF could be used at leaf nodes as in exhaustive ICTS; however, solving (small) branches as MAPFM instances using approaches other than branch-and-bound is also a direction to explore. To retain the benefits of ordered enumeration and to potentially allow for a higher degree of bounding, an optimized version \cite{miller1997} of Murty's algorithm \cite{murty1968} for ranking all matchings in increasing order of cost could be taken as a starting point for such a branch-and-bound approach.

>>>>

	\section{Introduction}
	Finding a shortest path between two vertices on a graph is a fundamental problem that appears in many contexts like GPS navigation \cite{sturtevant2010} and network routing \cite{broch1998}. In some applications like autonomous aircraft towing \cite{morris2016} and train unit routing in shunting yards \cite{mulderij2020}, planning agents separately can result in conflicts between agents as they follow their paths. To prevent such conflicts, agents have to be planned together to find a combination of non-conflicting paths. The problem of finding such a combination is the Multi-Agent Pathfinding (MAPF) problem \cite{stern2019}. In finding optimal solutions, one of two objectives is typically optimized: the makespan which is the cost of the longest path, or the Sum of Individual path Costs (SIC). Algorithms for optimally solving MAPF include Increasing Cost Tree Search \cite{sharon2013}, Conflict-Based Search \cite{sharon2015} and Branch-and-Cut-and-Price \cite{lam2019}.

@@

	\section{Multi-agent Pathfinding with Matching} % Problem description
	\label{mapfm}
	\begin{wrapfigure}[19]{r}{.25\textwidth}
		\begin{minipage}{\linewidth}
			\centering\captionsetup[subfigure]{justification=centering}
			\includegraphics[width=\linewidth]{img/vertex-conflict}
			\subcaption{Vertex conflict}
			\label{fig:conflictsa}\par\vfill
			\includegraphics[width=0.75\linewidth]{img/edge-conflict}
			\subcaption{Edge conflict}
			\label{fig:conflictsb}
		\end{minipage}
		\vspace{-10pt}
		\caption{Two types of conflict between agents (from Fig. 1 in \cite{stern2019})}\label{fig:conflicts}
	\end{wrapfigure}

@@

	\subsubsection{MDD generation for matching}
	To generate MDDs for MAPFM, each agent also has to consider not one goal but all matching goals. To do this, the MDD construction following the breadth-first search has to be changed: each goal becomes an endpoint from which paths to the start node are traced, resulting in an MDD with one start node and one or more terminal nodes.
	\begin{wrapfigure}{R}{.35\textwidth}
		\centering
		\includegraphics[width=\linewidth]{img/path}
		\caption{Pathological case for ICTS-m}
		\label{fig:path}
	\end{wrapfigure}
	\subsubsection{Pathological case}
	\label{pathology}
	As a consequence of the $\min$ operation in the root definition, a pathological case can be derived. Consider a team of 4 agents $a_1^1$ through $a_4^1$. Let them be arranged in a configuration as depicted in Figure \ref{fig:path}. All four agents match the color of the center goal and hence for each $i$, $c^*_i = 1$, so that the root is $(1,1,1,1)$ with SIC $4$. The optimal solution however is $(1,3,3,3)$ with SIC $10$ numbering the agents clockwise, meaning that $\Delta = 6$. It can be seen that instances of this form with arbitrarily large $\Delta$ can be constructed. As ICTS-m being a generalization of ICTS is $\mathcal{O}(k^\Delta)$, even just extending the limbs of the instance to yield an ICT root of $(1,9,9,9)$ with $\Delta = 24$ will result in an unfathomably large number of nodes being evaluated assuming no pruning or other modifications to the base algorithm.
	%\subsubsection{Inverted MDD search}
	%The uniqueness of goals guarantees the existence of a goal node without vertex conflicts in searching MDDs in MAPF. In MAPFM, this is not the case and causes joint MDDs to be searched fully without finding a result. In addition, even if there exists a goal node, many of the paths represented by the joint MDD lead to conflicting maximal-depth nodes.

@@

	\restoregeometry
	\subsection{Experimental setup}
	Because MAPFM lies on the fringe of pathfinding research, no standard benchmark exists. For MAPF, the Moving AI benchmarks \cite{sturtevant2012} are often used and potentially a new MAPFM standard benchmark could be defined by generating suitable team-partitions of agents for these MAPF instances; however, it was noticed that optimally solving even small MAPFM instances with comparatively few agents and short distances to goals was already a challenge computationally. This observation informed the creation of a map generator by the author of \cite{bruin2021} with various parameters.

@@

	\begin{enumerate}
		\item 25\% wall: a grid with 25\% of tiles being obstacles.
		\item 75\% wall: a grid with 75\% of tiles being obstacles. 
	\end{enumerate}
	For each of these types, instances were generated with between $1$ and $25$ agents either all in the same team or evenly partitioned into three teams. For each combination of map-type, number of teams and number of agents, 200 instances were generated. In scenarios with three teams, instances with $k < 3$ were not used. When solving the instances, for each combination of map-type and number of teams, the number of agents was increased until 0\% of instances could be solved within a timeout of 120s, at which point no more data was recorded for greater numbers of agents. This approach is similar to the method described in \cite{stern2019}.

	% For each of these types, maps were generated with $K = 1$ and $K = 3$ and $k$ from $K$ to $25$. When solving the instances, for each combination of map-type and $K$, $k$ was increased until 0\% of instances could be solved within a timeout of 120s, at which point no more data was recorded for larger values of $k$. This approach is similar to the method described in \cite{stern2019}.

@@

	\subsection{Experiment 1: ICTS-based MAPFM algorithms}
	To investigate the relative performance of various algorithms discussed in Section \ref{section:icts-matching}, a selection of both exhaustive and ICTS-m variants was made. Their performance in terms of fraction of problems solved within a timeout of 120s is shown in Figure \ref{fig:i-probs}. The meaning of the abbreviations used in the legend are as follows:
	\begin{itemize}
		\begin{minipage}{0.45\linewidth}
			\item S: Simple pruning
			\item E: Enhanced pruning
			\item C: Child pruning
		\end{minipage}
		\begin{minipage}{0.45\linewidth}
			\item ID: Independence detection
			\item B: Bounded search
			\item O: Approximate ordered enumeration
		\end{minipage}
	\end{itemize}
	Across scenarios, it is seen that typically exhaustive methods outperform ICTS-m algorithms (\ref{fig:i-probs}), with performance being more similar in the maps with 75\% wall. In particular on 75\% wall maps with three teams (\ref{fig:i-75-3-p}), the performance of ICTS-m+ID (and optional pruning) is seen to outperform bounded exhaustive search as $k$ increases, and some of these also manage to solve a small fraction of 75\% maps with one team that exhaustive methods can not within the time-out (\ref{fig:i-75-1-p}).

	Between exhaustive variants, with exception of the ID variant, a type of hierarchy can be distinguished, best illustrated by \ref{fig:i-25-3-p}; almost without exception, Exh+S+C+B+O outperforms Exh+E+B+O which outperforms Exh+E+B which in turn outperforms Exh+E. Bounded search outperforms unbounded search as expected, given that the former does as much work as the latter and often less. Approximate ordering outperforming unordered enumeration is not as obvious from a theoretical standpoint, as discussed in \ref{ordered-enum}, and simply indicates that ordered enumeration works well in practice. Simple pruning with child pruning outperforms exhaustive pruning but this needs to be qualified. Comparing Exh+S+B+O with Exh+E+B+O shows that part of the difference can be explained by the use of simple pruning instead of enhanced pruning. Nonetheless, in instances with 75\% wall (\ref{fig:i-75-1-p},\ref{fig:i-75-3-p}) there appears to be a real benefit to searching a smaller ICT. Across ICTS-m variants, similar trends can be observed but the differences across pruning strategies are too slight to draw any conclusions from. As discussed in Section \ref{future}, more work is needed to determine if such an improvement is to be expected more generally.

	The benefit of using ID together with ordered enumeration is less clear. Ordered enumeration with ID performs much better in \ref{fig:i-75-3-p}, a setting with more obstacles and more teams, but ordered enumeration without ID on the other hand is superior in an open setting with a single team (\ref{fig:i-25-1-p}). The overhead associated with the various planning and merging steps (see Algorithm \ref{id-algo}) could be the cause the slowdown observed once $M$ grows large (\ref{fig:i-25-1-p},\ref{fig:i-75-1-p}); once the upper bound is tightened, ICTs only have to be searched shallowly so there is no point in going through the steps of ID. This explains why no speedup is seen, as between ICTS-m and ICTS-m+ID. Another factor is that if all agents belong to the same team, it is unlikely that there will be a significant benefit to using ID, as agents from the same team planned independently often route to the same goal.

@@

	CBM solving all problems with one team (Figure \ref{fig:r-25-1-p} and \ref{fig:r-75-1-p}), problems equivalent to anonymous MAPF instances, is at first surprising but follows directly from the definition of the CBM method. CBM solves single teams as anonymous MAPF instances in polynomial time, meaning that solving scenarios with a single team reduces to a polynomial-time procedure, namely SSP. Following the results reported in \cite{ma2016}, it is expected that as the number of teams approaches $k$, CBM performance will deteriorate and thus become more similar to the results for exhaustive methods here reported.
	%	Still, some implementations might be more optimized than others so this might be a factor in performance characteristics and in particular in relative performance differences.
	\section{Conclusions and Future work}
	\label{conclusions}
	\label{future}
	The goal of this research was to seek methods for solving MAPFM based on ICTS. In this, the research has succeeded, but the scalability of the algorithms leaves something to be desired. Exhaustive methods, like those discussed in detail in Subsection \ref{exhaustive}, cannot be expected to scale well as teams grow larger due to their complexity, no matter what optimizations are put in place. Nonetheless, as was shown in Section \ref{experiments}, the more advanced exhaustive ICTS algorithms consistently outperform the ICTS-m based algorithms. Compared to the SIC-adaptation of the CBM method \cite{baauw2021}, it was seen that exhaustive methods generally perform relatively poorly if there are fewer teams and fewer obstacles, while being competitive once teams grow smaller and maps become more constrained. Delineating under which conditions exhaustive methods outperform CBM and vice versa is something for further research.

	%\section{Future work}
	A key idea for future research to look into is how to exploit the lower-bound on solution SIC that can be found using the Hungarian Algorithm \cite{kuhn55} or similar methods like the JVC algorithm \cite{jonker1987}. This lower-bound could be used in a branch-and-bound \cite{land1960} search of matchings. An algorithm for MAPF could be used at leaf nodes as in exhaustive ICTS; however, solving (small) branches as MAPFM instances using approaches other than branch-and-bound is also a direction to explore. To retain the benefits of ordered enumeration and to potentially allow for a higher degree of bounding, an optimized version \cite{miller1997} of Murty's algorithm \cite{murty1968} for ranking all matchings in increasing order of cost could be taken as a starting point for such a branch-and-bound approach.
